% ==========================================================================
% main.tex  —  Standard Article Format (A4, 12pt)
% Title     : fMRI Brain Data Analysis using SPM and MATLAB
% Author    : Fahim Ur Rehman Shah (EE2629)
% Course    : CSE532 — Signal and Image Processing (MS Level)
% Supervisor: Dr. Adnan Shah, FCSE, GIKI
% Date      : May 2026
% Compile   : pdflatex main.tex  (run twice)
%             bibtex main
%             pdflatex main.tex  (twice more)
% ==========================================================================

\documentclass[12pt,a4paper]{article}

% --- Page layout ---
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}

% --- Essential packages ---
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{float}          % [H] placement — figures stay exactly where written
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{textcomp}
\usepackage[hidelinks]{hyperref}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{cite}

% --- Caption styling ---
\usepackage[font=small, labelfont=bf, margin=0.5cm]{caption}

% --- Header / footer with page numbers ---
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small fMRI Analysis using SPM and MATLAB}
\fancyhead[R]{\small Fahim Ur Rehman Shah (EE2629)}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}

% --- Section title formatting ---
\usepackage{titlesec}
\titleformat{\section}{\large\bfseries}{\thesection.}{0.5em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{0.5em}{}

% --- Line spacing ---
\usepackage{setspace}
\setstretch{1.15}

\graphicspath{{figures/}}

% ==========================================================================
\begin{document}

% --------------------------------------------------------------------------
% Title block
% --------------------------------------------------------------------------
\begin{center}
  {\LARGE\bfseries fMRI Brain Data Analysis using SPM and MATLAB:\\[0.3em]
  Preprocessing, GLM, Statistical Mapping and Functional Connectivity}

  \vspace{1em}

  {\large Fahim Ur Rehman Shah}\\[0.2em]
  \textit{Faculty of Computer Science \& Engineering (FCSE)}\\
  \textit{GIK Institute of Engineering Sciences \& Technology}\\
  Registration No.: EE2629 \quad|\quad \texttt{gee2629@giki.edu.pk}

  \vspace{0.5em}

  {\large Supervisor: Dr.\ Adnan Shah}\\[0.2em]
  \textit{Assistant Professor, FCSE, GIK Institute}\\
  CSE532 --- Signal and Image Processing

  \vspace{0.5em}
  \textit{May 2026}
\end{center}

\vspace{0.5em}
\hrule
\vspace{1em}

% --------------------------------------------------------------------------
% Abstract + Keywords
% --------------------------------------------------------------------------
\begin{abstract}
\input{sections/abstract}
\end{abstract}

\noindent\textbf{Keywords:} fMRI, SPM25, preprocessing, general linear model,
BOLD signal, functional connectivity, motor cortex, ventral visual stream,
OpenNeuro, MATLAB

\vspace{1em}
\hrule
\vspace{1em}

% ==========================================================================
\section{Introduction}
\input{sections/introduction}

% ==========================================================================
\section{Datasets and Experimental Design}
\label{sec:datasets}

Two publicly available datasets from the OpenNeuro repository were
used in this project, chosen to cover distinct functional systems.

\subsection{Dataset 1 — ds000114 (Motor Task)}
The \textit{test-retest fMRI dataset}~\cite{Gorgolewski2013}
(ds000114) includes a blocked motor task in which subjects perform
repetitive finger tapping, foot tapping, and lip movements, each
interleaved with rest. Subject \texttt{sub-01}, session
\texttt{ses-test}, run~1 was analysed (184 volumes). The T1
anatomical image from the same session was used for normalisation.
Scan parameters were read directly from the NIfTI header.

\subsection{Dataset 2 — ds000105 (Visual Task)}
The \textit{visual object recognition} dataset~\cite{Haxby2001}
(ds000105) presents eight visual categories — faces, houses, cats,
bottles, scissors, shoes, chairs, and scrambled images — in a
blocked design. Subject \texttt{sub-1}, run~1 was analysed (121
volumes). This dataset is widely used as a benchmark for multivariate
pattern analysis; running a standard GLM on it provides an immediate
sanity check against published results.

% ==========================================================================
\section{Methods}
\label{sec:methods}
\input{sections/methods}

% ==========================================================================
\section{Results}
\label{sec:results}
\input{sections/results}

% ==========================================================================
\section{Discussion}
\label{sec:discussion}

The results across both datasets were broadly consistent with
expectations from the neuroimaging literature, which suggests the
pipeline is working correctly.

\textbf{Motion quality.}
Mean framewise displacement was below 0.2\,mm in both scans, putting
them in the ``low motion'' category under the criteria of Power
et~al.~\cite{Power2012}. Clean motion profiles make it unlikely that
the activation maps are motion artefacts.

\textbf{Motor cortex somatotopy.}
The three finger/foot/lips contrasts in Dataset~1 produced activation
peaks that followed the expected somatotopic organisation of primary
motor cortex from medial (foot) to lateral (face/lips)
territory~\cite{Lotze2000}. The omnibus AllMotor map covered the
entire motor strip plus the cerebellum and thalamus.

\textbf{Visual category selectivity.}
The face-selective fusiform cluster in Dataset~2 (Fig.~\ref{fig:ds2:tmap:face})
matches the reported location of the fusiform face area in the
original Haxby et~al.\ paper~\cite{Haxby2001}. The other category
contrasts produced distributed activation in the lateral
occipito-temporal cortex, consistent with object-selective processing
in the ventral stream.

\textbf{Functional connectivity.}
Both connectivity maps showed spatially coherent networks. The motor
seed produced bilateral M1 and SMA connectivity, and the visual seed
produced occipito-temporal connectivity. The cross-dataset comparison
(Fig.~\ref{fig:comparison:connectivity}) confirms these networks are
spatially non-overlapping. Negative correlations with default-mode
regions are also expected~\cite{Biswal1995}.

\textbf{Limitations.}
This analysis covers a single subject per dataset. The connectivity
analysis used the task run rather than a resting-state scan, so
correlations partly reflect task-related coactivation. Future work
should include group-level analyses and separate resting-state data.

% ==========================================================================
\section{Conclusion}
\label{sec:conclusion}
\input{sections/conclusion}

% ==========================================================================
\section*{Acknowledgements}
The author thanks Dr.\ Adnan Shah (FCSE, GIKI) for supervision and
guidance throughout this project. The OpenNeuro platform and the
contributors of datasets ds000114 and ds000105 are gratefully
acknowledged.

% ==========================================================================
\bibliographystyle{IEEEtran}
\bibliography{bibliography}

\end{document}
